On the presentation of orthogonal groups over polynomial rings

### 2019-05-23 dl-paper.tex
Fixed some issues

\begin{comment}
Suppose for a moment that $\langle \alpha, \beta \rangle = -1$ and  $\langle \beta, \alpha \rangle = -1$ then
\[ \{s, t^{-1} \} = \{s,  t^{-1}\}_\alpha = \{t, s^{-1} \}_\beta^{-1} = \{s^{-1}, t\} \]
In particular, $\{s, s^{-1}\} = \{s, s^{-1}\}^{-1}$ 
\end{comment}
\subsection{Unstable Quillen K-groups} \label{sec:quillen}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following reqirements:
\begin{enumerate}
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coequalizers of kernel pairs.
\end{enumerate}

@@

\subsection{The subgroups $P_\alpha(0)$, $P_\alpha(*)$ and their properties.} 
Throughout this subsection $\Phi$ is a simply-laced root system of rank $\geq 3$ and $\alpha$ denotes some fixed root of $\Phi$.

Consider the following root subsets of $\Phi$ (cf.~\cite[Proposition~5.7]{St71}):
\begin{equation} \nonumber \begin{align}
   Z_+(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle > 0 \}, \\
   Z_0(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle = 0 \}, \\
   Z(\alpha)   & = Z_0(\alpha) \sqcup Z_+(\alpha). \end{align} \end{equation}

We denote by $Z_\alpha(A, M)$ the subgroup of $\overline{\St}(\Phi, A, M)$ generated by elements
 $x_{\beta}(m),\ \beta \in Z_+(\alpha)$ and $z_{\gamma}(m, \zeta),\ \gamma \in Z_0(\alpha),$ where $m \in M$, $\zeta \in A$.
It is not hard to see that \[Z_\alpha(A, M) = \Img\left(\overline{\St}(Z_0(\alpha), A, M) \to \overline{\St}(\Phi, A, M)\right) \rtimes \hat{U}(Z_+(\alpha), M). \]
Notice that the subgroup $Z_\alpha(A, A)$ coincides with $Z(\hat{U}_\alpha)$ defined in~\cite[984]{St71}.

\begin{rem}\label{Z-DS} Notice that our assumptions on the rank of $\Phi$ guarantee that $Z_0(\alpha)$ is nonempty, 
hence $Z_\alpha(A, M)$ contains relative Dennis--Stein symbols $\langle a, m \rangle$ for $a\in A$, $m\in M$.
By~\eqref{DS-S-relationship} relative Steinberg symbols $\{a, 1+m\}$ are also contained in $Z_\alpha(A, M)$ for all $a\in A^*$, $m\in M$. \end{rem}

@@

Suppose $g$ has type 1 or 2 and $\beta - \alpha \in \Phi$. By~\cref{Zrels} we obtain that
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mX^2f) \cdot z_{\beta}(Xf, \xi), \label{eq3-1} \\
  z_{\beta}(f, X\xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi ) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mXf) \cdot z_{\beta}(f, X\xi), \label{eq3-2} \end{align}
The expressions in the right hand sides of~\eqref{eq3-1} and~\eqref{eq3-2} are products of generators of type 4, 1, 1 and 4, 1, 2, respectively.

@@

\begin{rem} \label{rem:palpha} From the two preceding lemmas it follows that the value of $p_\alpha$ on an element $g\in K(A[X], M[X])$ can be computed via the following procedure. 
Start with any presentation of $g$ as a product of elements $z_{\beta}(Xf, \xi)$ for $\beta \in \Phi$ and $c_\delta(f, X\xi)$ for some fixed $\delta\in Z_0(\alpha)$. 
Now pick among these factors those that correspond to the root $\beta = -\alpha$ (i.\,e. pick all factors $z_{-\alpha}(Xf_i, \xi_i)$). 
Now $p_\alpha(g)$ is precisely the sum of constant terms of the polynomials $f_i$. \end{rem}

@@

\begin{lemma} Let $\Phi$ be any root system of rank $\geq 3$, $\alpha \in \Phi$ and $(A, M)$ be a local pair.
 Then the partial function $t_\alpha(aX^{-1}) \colon V \not \to V$ descends to a well-defined total function 
  $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $(g, h, u) \in V$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev^*_{X=0}(g_1)^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$,
  which shows that $t_\alpha(aX^{-1})$ is defined on $(g, h \cdot i_+(h_0), u)$, while the latter triple also lies in the same orbit as $(g, h, u)$.

 Next, let us show that the value of $T_\alpha(aX^{-1})$ does not depend on the choice of a representative.
 Let $(g, h_1, u) \sim (g, h_2, u)$ be elements of the same orbit for which both $p(g, h_1, u)$ and $p(g, h_2, u)$ belong to $P_\alpha(*)$.
 By definition, $h_1^{-1} h_2 = i_-(h_0)$, for some $h_0 \in G^0_M$, moreover, 
  $p(g, h_1, u)^{-1} \cdot p(g, h_2, u) = j_- i_-(h_0) = j_+ i_+(h_0) \in P_\alpha(*)$.
 By~\cref{Pstar-char} there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $g_0 := ev^*_{X=0}(g_1) \in Z_\alpha(A, M)$ and $j_+(g_1) = j_+ i_+ (h_0)$.
 From the last equality and the injectivity of the map $G(\Phi, A[X]) \to G(\Phi, A[X, X^{-1}])$ we obtain that
 the projections of $g_0$, $g_1$ and $h_0$ in $G(\Phi, A[X, X^{-1}])$ are equal, which shows that $g_0 \cdot h_0^{-1} \in \overline{\K_2}(\Phi, A, M)$. 
 By~\cite[Theorem~2.5]{Ste73} the latter subgroup is generated by relative Steinberg symbols $\{ a, 1 + m \}_\beta$, where $\beta$ is any fixed long root of $\Phi$,
 It is not hard to check using e.\,g. the table on p.~986 of~\cite{St71} that $Z_0(\alpha)$ always contains a long root, therefore
  $\overline{\K_2}(\Phi, A, M)$ is contained in $Z_\alpha(A, M)$ by~\cref{Z-DS} hence $h_0 \in Z_\alpha(A, M)$.
 Thus, we have shown that $i_-(h_0)$ is centralized by $X_\alpha(A[X^{-1}])$, which allows us to conclude that
 \begin{multline} \nonumber
  t_\alpha(aX^{-1})(g, h_1, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h_1 \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right) \sim \\
  \sim \left( x_\alpha(aX^{-1})\cdot g ,\ h_1 i_-(h_0) \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right) = t_\alpha(aX^{-1})(g, h_2, u). \qedhere
 \end{multline} \end{proof}

>>>>

\begin{comment}
Suppose for a moment that $\langle \alpha, \beta \rangle = -1$ and  $\langle \beta, \alpha \rangle = -1$ then
\[ \{s, t^{-1} \} = \{s,  t^{-1}\}_\alpha = \{t, s^{-1} \}_\beta^{-1} = \{s^{-1}, t\} \]
In particular, $\{s, s^{-1}\} = \{s, s^{-1}\}^{-1}$ 
\end{comment}

\subsection{Unstable Quillen K-groups} \label{sec:quillen}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following reqirements:
\begin{enumerate}
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coequalizers of kernel pairs.
\end{enumerate}

@@

\subsection{The subgroups $P_\alpha(0)$, $P_\alpha(*)$ and their properties.} 
Assume for a moment that $\Phi$ is an arbitrary root system and consider the following subsets of $\Phi$:
\begin{equation} \nonumber \begin{align}
   Z_+(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle > 0 \}, \\
   Z_0(\alpha) & = \{ \beta \in \Phi \mid \alpha + \beta \not\in \Phi,\ \langle \alpha, \beta \rangle = 0 \}, \\
   Z(\alpha)   & = Z_0(\alpha) \sqcup Z_+(\alpha). \end{align} \end{equation}
Clearly, $Z_0(\alpha)$ (resp. $Z_+(\alpha)$) is a reductive (resp. special) subset of $\Phi$, moreover, in the simply-laced case the assumption $\alpha+\beta\not\in \Phi$ in the definition of $Z_0(\alpha)$ is superfluous, i.\,e. $Z_0(\alpha) = \{ \alpha\in\Phi \mid \alpha \perp \beta \}$
(cf.~\cite[Proposition~5.7]{St71}).

We denote by $Z_\alpha(A, M)$ the subgroup of $\overline{\St}(\Phi, A, M)$ generated by elements
 $x_{\beta}(m),\ \beta \in Z_+(\alpha)$ and $z_{\gamma}(m, \zeta),\ \gamma \in Z_0(\alpha),$ where $m \in M$, $\zeta \in A$.
It is not hard to see that \[Z_\alpha(A, M) = \Img\left(\overline{\St}(Z_0(\alpha), A, M) \to \overline{\St}(\Phi, A, M)\right) \rtimes \hat{U}(Z_+(\alpha), M). \]
It is clear, that $Z_\alpha(A, M)$ centralizes the root subgroup $X_\alpha(A)$ (cf. \cite[984]{St71}).

For the rest of this subsection $\Phi$ is a simply-laced root system of rank $\geq 3$ and $\alpha$ is a root of $\Phi$.

\begin{rem}\label{Z-DS} Notice that our assumptions on the rank of $\Phi$ guarantee that $Z_0(\alpha)$ is nonempty, hence the group $Z_\alpha(A, M)$ contains relative Dennis--Stein symbols $\langle a, m \rangle$ for $a\in A$, $m\in M$.
By~\eqref{DS-S-relationship} relative Steinberg symbols $\{a, 1+m\}$ are also contained in $Z_\alpha(A, M)$ for all $a\in A^*$, $m\in M$. \end{rem}

@@

Suppose $g$ has type 1 or 2 and $\alpha - \beta \in \Phi$. By~\cref{Zrels} we obtain that
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mX^2f) \cdot z_{\beta}(Xf, \xi), \label{eq3-1} \\
  z_{\beta}(f, X\xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi ) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mXf) \cdot z_{\beta}(f, X\xi), \label{eq3-2} \end{align}
The expressions in the right hand sides of~\eqref{eq3-1} and~\eqref{eq3-2} are products of generators of type 4, 1, 1 and 4, 1, 2, respectively.

@@

\begin{rem} \label{rem:palpha} From the two preceding lemmas it follows that the value of the function $p_\alpha$ from the statement of~\cref{P0_normal} on an element $g\in K(A[X], M[X])$ can be computed via the following procedure. 
Start with any presentation of $g$ as a product of elements $z_{\beta}(Xf, \xi)$ for $\beta \in \Phi$ and $c_\delta(f, X\xi)$ for some fixed $\delta\in Z_0(\alpha)$. 
Now pick among these factors those that correspond to the root $\beta = -\alpha$ (i.\,e. pick all factors $z_{-\alpha}(Xf_i, \xi_i)$). 
Now $p_\alpha(g)$ is precisely the sum of constant terms of the polynomials $f_i$. \end{rem}

@@

\begin{lemma} Let $\Phi$ be a simply-laced root system of rank $\geq 3$, $(A, M)$ be a local pair. Then for $\alpha \in \Phi$ the partial function $t_\alpha(aX^{-1}) \colon V \not \to V$ descends to a well-defined total function 
  $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $(g, h, u) \in V$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev^*_{X=0}(g_1)^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$,
  which shows that $t_\alpha(aX^{-1})$ is defined on $(g, h \cdot i_+(h_0), u)$, which is also a representative of the same orbit as $(g, h, u)$.

 Next, let us show that the value of $T_\alpha(aX^{-1})$ does not depend on the choice of a representative.
 Let $(g, h_1, u) \sim (g, h_2, u)$ be elements of the same orbit for which both $p(g, h_1, u)$ and $p(g, h_2, u)$ belong to $P_\alpha(*)$.
 By definition, $h_1^{-1} h_2 = i_-(h_0)$, for some $h_0 \in G^0_M$, moreover, 
  $p(g, h_1, u)^{-1} \cdot p(g, h_2, u) = j_- i_-(h_0) = j_+ i_+(h_0) \in P_\alpha(*)$.
 By~\cref{Pstar-char} there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $g_0 := ev^*_{X=0}(g_1) \in Z_\alpha(A, M)$ and $j_+(g_1) = j_+ i_+ (h_0)$.
 From the last equality and the injectivity of the map $G(\Phi, A[X]) \to G(\Phi, A[X, X^{-1}])$ we obtain that
 the projections of $g_0$, $g_1$ and $h_0$ in $G(\Phi, A[X, X^{-1}])$ are equal, which shows that $g_0 \cdot h_0^{-1} \in \overline{\K_2}(\Phi, A, M)$. 
 By~\cref{thmStein} the latter subgroup is generated by relative Steinberg symbols $\{ a, 1 + m \}$ and therefore is contained in $Z_\alpha(A, M)$ by~\cref{Z-DS}. Thus, we have obtained that $h_0 \in Z_\alpha(A, M)$ and hence that $i_-(h_0)$ is centralized by $X_\alpha(A[X^{-1}])$, which allows us to conclude that
 \begin{multline} \nonumber
  t_\alpha(aX^{-1})(g, h_1, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h_1 \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right) \sim \\
  \sim \left( x_\alpha(aX^{-1})\cdot g ,\ h_1 i_-(h_0) \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right) = t_\alpha(aX^{-1})(g, h_2, u). \qedhere
 \end{multline} \end{proof}
